Enzo: An Adaptive Mesh Refinement Code for Astrophysics

### 2013-05-28 numerical-cooling.tex
Added most of Dave Collins' suggested edits (except for a bunch of stuff having to do with the hydro).
We definitely use Glover & Abel 2008 by default for H2, and the code suggests we use Lepp & Shull as well as GA98 for HD.

\enzo\ has multiple methods for computing the energy change from
radiative cooling and heating.  All of them assume the gas is totally
optically thin.  Below we describe the methods for computing the
cooling rates from metal-free and metal-enriched gas.  Sample cooling
curves for each of \enzo's primary cooling methods are shown in Figure
\ref{fig.cooling_rate}.

@@

As discussed in Section~\ref{sec.num.chemistry}, the set of reactions that
characterize a metal-free gas is simple enough to be computed in
non-equilibrium during the simulation.  Similarly, the radiative
cooling of metal-free gas is solved by directly computing the cooling
and heating rates from the following individual processes for atomic H
and He: collisional excitation and ionization, recombination,
free-free emission, Compton scattering off of the cosmic microwave
background (CMB), and photo-heating from a UV metagalactic background.
If the H$_{2}$ chemistry network is enabled, the following H$_{2}$
cooling processes are also considered: ro-vibrational transitions
\citep{1998A&A...335..403G}, heating and cooling from molecular
formation and destruction \citep{2009Sci...325..601T}, and
collision-induced emission \citep{2004MNRAS.348.1019R}.  If Deuterium
chemistry is enabled, then rotational transitions of HD
\citep{1998A&A...335..403G} are treated as well.  The radiative
cooling calculation is coupled to the update of the chemistry network
such that they both occur within the same subcycling loop.  This is
necessary in regimes where rapid cooling or change in the ionization
state occur.  In addition to the subcycle timestepping contraints
mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep is also not
permitted to exceed 10\% of the cooling time, $t_{cool} = e/\dot{e}$.  A
metagalactic background affects the gas through both photo-heating and
photo-ionization.  These are treated by including redshift-dependent
photo-ionization and photo-heating rate terms in the chemistry and
cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.  The
black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both
with and without a radiation background.  More detail on the specific
UV backgrounds present in \enzo is presented below in
Section~\ref{sec.num.rad-homogeneous}.

>>>>

\enzo\ has multiple methods for computing the energy change from radiative
cooling and heating.  All of them assume the gas can be modeled either as
completely optically thin or with a simple, local approximation to optical
thickness.  Below we describe the methods for computing the cooling rates from
metal-free and metal-enriched gas.  Sample cooling curves for each of \enzo's
primary cooling methods are shown in Figure \ref{fig.cooling_rate}.

@@

As discussed in Section~\ref{sec.num.chemistry}, the set of reactions that
characterize a metal-free gas is simple enough to be computed in
non-equilibrium during even very simulations.  Similarly, the radiative
cooling of metal-free gas is solved by directly computing the cooling
and heating rates from the following individual processes for atomic H
and He: collisional excitation and ionization, recombination,
free-free emission, Compton scattering off of the cosmic microwave
background (CMB), and photo-heating from a UV metagalactic background.
If the H$_{2}$ chemistry network is enabled, the following H$_{2}$
cooling processes are also considered: ro-vibrational transitions
\citep{2008MNRAS.388.1627G,1998A&A...335..403G}, heating and cooling from molecular
formation and destruction \citep{2009Sci...325..601T}, and
collision-induced emission \citep{2004MNRAS.348.1019R}.  If Deuterium
chemistry is enabled, then rotational transitions of HD
\citep{1998A&A...335..403G, 1983ApJ...270..578L} are treated as well.  The
radiative cooling calculation is coupled to the update of the chemistry network
such that they both occur within the same subcycling loop.  This is necessary
in regimes where rapid cooling or change in the ionization state occur, as this
will influence the kinetic rate coefficients through both changes in energy and
the equation of state of the gas.  In addition to the subcycle timestepping
contraints mentioned in Section~\ref{sec.num.chemistry}, the subcycle timestep
is also not permitted to exceed 10\% of the cooling time, $t_{cool} =
e/\dot{e}$.  A metagalactic background affects the gas through both
photo-heating and photo-ionization.  These are treated by including
redshift-dependent photo-ionization and photo-heating rate terms in the
chemistry and cooling equations for \ion{H}{1}, \ion{He}{1}, and \ion{He}{2}.
The black curves in Figure \ref{fig.cooling_rate} show cooling rates for a
metal-free gas with number density $n_{\rm H} = 10^{-4}$ cm$^{-3}$ both with
and without a radiation background.  More detail on the specific UV backgrounds
present in \enzo\ is presented below in Section~\ref{sec.num.rad-homogeneous}.
